Distributional Trees and Forests

### 2017-10-06 pkg/disttree/inst/draft/draft.Rnw
added prediction on one station with 27 years as learning data and 3 years as testing data
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1629 255f4249-2263-4417-b1ea-8498779e27a7

\begin{tabular}{| l | c |}
\hline
basic covariates & variations
\hline
tppow: total precipitation power transformed & tppow_mean, tppow_spread, tppow_min, tppow_max \\
& tppow_0612, tppow_1218, tppow_1824, tppow_2430 \\
\hline
capepow:    & capepow_mean, capepow_spread, capepow_min, capepow_max \\
& capepow_0612, capepow_1218, capepow_1824, capepow_2430 \\
\hline
tcloc: total cloud coverage & \\
\hline
\hline
\end{table}

@@

\begin{tabular}{| l | c  c  c  c |}
\hline
model & mu link & sigma link & mu covariates & sigma covariates \\
\hline
disttree & identity & log & all covariates & all covariates \\ 
\hline
distforest & identity & log & all covariates & all covariates \\
\hline
gamlss & identity & log & manually selected covariates: & manually selected covariates \\
\hline
gamboostLSS & identity & log & all covariates & all covariates \\
\hline
EMOS & identity & log & tppow_mean & tppow_spread \\
\hline
\end{table}

These methods are compared in a 10x10 cross validation framework based on the resulting RMSE, log-likelihood and CRPS.\\
These values are illustrated in the following boxplots.

<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("rain_Axams.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_Axams.rda")
  load("rain_Axams.rda")
} else {
  #source("~/svn/partykit/pkg/disttree/inst/draft/rain_cross.R")
  source("rain_cross.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  setwd("~/svn/partykit/pkg/disttree/inst/draft/")
  rainres <- rain_cross(stationname = "Axams", 
                        seedconst = 7, ntree = 100,
                        forest_mtry = 27,
                        gamboost_cvr = TRUE,
                        frac = FALSE)
}

>>>>

\begin{tabular}{| l | c | l |}
\hline
basic covariates & nr & variations\\
\hline
\textbf{tp{\_}pow}: total precipitation & 12 & mean, std dev, min, max (all for 6--30 UTC)\\
\hspace*{\fill} (power transformed by 1.6) & & mean: 6--12, 12--18, 12--24, 24--30 UTC\\
& & std dev: 6--12, 12--18, 12--24, 24--30 UTC\\
\hline
\textbf{cape{\_}pow}: convective available & 12 & mean, std dev, min, max (all for 6--30 UTC)\\
\hspace*{\fill} potential energy & & mean: 6--12, 12--18, 12--24, 24--30 UTC\\
& & std dev: 6--12, 12--18, 12--24, 24--30 UTC\\
\hline
dswrf: downwards short & 4-6 & mean, std dev, min, max (all for 0--30 UTC)\\
\hspace*{\fill}ware radiation flux (sunshine) & & mean of the spread, max of the spread, (min of the spread)\\
\hline
msl: mean sea level pressure & 6 &
mean of the mean, max of the mean, min of the mean,
mean of the spread, max of the spread, min of the spread \\
\hline
pwat: preciptable water & 6 &
mean of the mean, max of the mean, min of the mean, 
mean of the spread, max of the spread, min of the spread \\
\hline
tmax: 2m maximum temperature & 6 &
mean of the mean, max of the mean, min of the mean, 
mean of the spread, max of the spread, min of the spread \\
\hline
tcolc: Total column-integrated condensate & 6 &
mean of the mean, max of the mean, min of the mean, 
mean of the spread, max of the spread, min of the spread \\
\hline
t500: temperature on 500 hPa & 6 &
mean of the mean, max of the mean, min of the mean, 
mean of the spread, max of the spread, min of the spread \\
\hline
t700: temperature on 700 hPa & 6 &
mean of the mean, max of the mean, min of the mean, 
mean of the spread, max of the spread, min of the spread \\
\hline
t850: temperature on 850 hPa & 6 &
mean of the mean, max of the mean, min of the mean, 
mean of the spread, max of the spread, min of the spread \\
\hline
tdiff500850: temperature difference & 3 &
mean, min, max \\
\hline
tdiff500700: temperature difference & 3 &
mean, min, max \\
\hline
tdiff700850: temperature difference & 3 &
mean, min, max \\
\hline
msl{\_}diff: temperature difference & 1 & 
- \\
\hline
\end{tabular}

@@

\begin{tabular}{| l | c  c  c  c |}
\hline
model & mu link & sigma link & mu covariates & sigma covariates \\
\hline
disttree & identity & log & all covariates & all covariates \\ 
\hline
distforest & identity & log & all covariates & all covariates \\
\hline
gamlss & identity & log & manually selected covariates: & manually selected covariates \\
\hline
gamboostLSS & identity & log & all covariates & all covariates \\
\hline
EMOS & identity & log & tppowmean & tppowspread \\
\hline
\end{tabular}

\subsection{Station Axams}
<<echo=FALSE, results=hide>>=
#### Axams prediction 27 - 3
if(file.exists("rain_Axams_pred.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_Axams_pred.rda")
  load("rain_Axams_pred.rda")
} else {
  #source("~/svn/partykit/pkg/disttree/inst/draft/rain_axams_pred.R")
  source("rain_axams_pred.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  res <- rain_axams_pred(seedconst = 7, ntree = 100,
                         forest_mtry = 27,   # if frac == TRUE: nvar/3 = 30
                         gamboost_cvr = FALSE,
                         frac = FALSE)

  save(res, file = "rain_Axams_pred.rda")
}

# prepare data for plot of estimated density functions
pday <- 24  # 2 (hohe Beobachtung zu niedrig geschaetzt), 4, evtl. auch 7, 8, 23 (eine 0-Beobachtung und 2 sehr aehnliche),

pdays <- if(pday<19) c(pday, pday + 31, pday + 61) else c(pday, pday + 30, pday + 61)
pdf <- predict(res$df, newdata = res$testdata[pdays,], type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))
cbind(df_exp, res$testdata[pdays,"robs"])

# plot predicted distributions together with observations
#set.seed(res$call$seedconst)
set.seed(7)
x <- sort(runif(500,-2,12))
y1 <- crch::dcnorm(x, mean = df_mu[1], sd = df_sigma[1], left = 0)
y2 <- crch::dcnorm(x, mean = df_mu[2], sd = df_sigma[2], left = 0)
y3 <- crch::dcnorm(x, mean = df_mu[3], sd = df_sigma[3], left = 0)
dayending <- if(pday > 3) "th" else switch(pday, "1" = {dayending <- "st"}, "2" = {dayending <- "nd"}, "3" = {dayending <- "rd"})

\begin{figure}[t!]
<<rain_Axams_pred, fig=TRUE, echo=FALSE>>=
plot(x = x, y = y1, type = "l", col = "red", 
         main = paste0("July ", pday, dayending), ylab = "density", ylim = c(0,max(y1, y2, y3) + 0.01))
    lines(x = x, y = y2, type = "l", col = "blue")
    lines(x = x, y = y3, type = "l", col = "darkgreen")
    legend('topright', c("2010", "2011", "2012"), col = c("red", "blue", "darkgreen"), lty = 1, cex = 1)

    p1 <- c(res$testdata[pdays,"robs"][1], crch::dcnorm(res$testdata[pdays,"robs"][1], mean = df_mu[1], sd = df_sigma[1], left = 0))
    p2 <- c(res$testdata[pdays,"robs"][2], crch::dcnorm(res$testdata[pdays,"robs"][2], mean = df_mu[2], sd = df_sigma[2], left = 0))
    p3 <- c(res$testdata[pdays,"robs"][3], crch::dcnorm(res$testdata[pdays,"robs"][3], mean = df_mu[3], sd = df_sigma[3], left = 0))

    points(x = p1[1], y = p1[2], col = "red")
    points(x = p2[1], y = p2[2], col = "blue")
    points(x = p3[1], y = p3[2], col = "darkgreen")

    lines(x = c(p1[1], p1[1]), y = c(p1[2], 0), col = "darkgrey", type = "l", lty = 2)
    lines(x = c(p2[1], p2[1]), y = c(p2[2], 0), col = "darkgrey", type = "l", lty = 2)
    lines(x = c(p3[1], p3[1]), y = c(p3[2], 0), col = "darkgrey", type = "l", lty = 2)
@

\caption{\label{}}
\end{figure}

The methods are compared in a 10x10 cross validation framework based on the resulting RMSE, log-likelihood and CRPS.\\
These values are illustrated in the following boxplots.

<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("rain_Axams.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_Axams.rda")
  load("rain_Axams.rda")
} else {
  #source("~/svn/partykit/pkg/disttree/inst/draft/rain_cross.R")
  source("rain_cross.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  rainres <- rain_cross(stationname = "Axams", 
                        seedconst = 7, ntree = 100,
                        forest_mtry = 27,
                        gamboost_cvr = TRUE,
                        frac = FALSE)

  save(rainres, file = "rain_Axams.rda")
}
